### 2017-02-07 abstract.tex
first

In this work we address the problem of finding unsafe behaviors with
given a safety property for a hybrid dynamical system. We treat the
system as a black box and approach the problem in two distinct steps.
First, we model the behavior of the system using a piecewise affine
discrete time model. Such a model is incomplete in nature and is
computed with respect to the given property. Next, we encode the
search for falsification as a bounded model checking query and use an
SMT solver to find a counter example. If found, we check the existance
of the violation in the original system to ensure reproducibility.

%Time bounded LTL property in 
%Models from verificaiton
%model without data
%black box model/data -> structured model