### 2017-04-07 abstract.tex
.

% Sampling based falsification techniques for black box hybrid dynamical
% systems use numerical simulations to search for violations of
% properties (specified in temporal logic). This constraints them to a
% `best effort' category unlike rigorous verification techniques, which
% however require a white box model. This work tries to bridge the gap
% by presenting an automatic falsification technique which models the
% black box system before searching for a violation. It proceeds by (a)
% modeling the behavior of the system using a piecewise affine (PWA)
% discrete time model and, (b) encoding the search for a violation as a
% standard bounded model checking problem. The resulting problem is a
% disjunction of linear programs and can be solved directly using
% off-the-shelf SMT solvers or by multiple runs of a linear programming
% solver. If a counter-example is found, the existence of a
% corresponding violation in the original system is checked for
% reproducibility.

Models of embedded control systems are often so complex that they can
be treated as effectively black-box systems. Sampling-based methods
for testing such systems use numerical simulations to search for
violations of properties (specified in a suitable requirement
formalism, such as temporal logic). This constrains them to a `best
effort' category unlike rigorous verification techniques, which
however require a model amenable to symbolic analysis. This work tries
to bridge the gap by presenting an automatic falsification technique
which learns a data-driven abstraction of the black box system before
searching for a violation. It proceeds by (a) modeling the behavior of
the system using a piecewise affine (PWA) discrete-time {\em
relational} model and, (b) encoding the search for a violation in the
abstract model as a bounded model checking problem. The bounded model
checking query in this instance can be viewed as a disjunction of
conjunctions of linear constraints (i.e. linear programs) and can be
solved directly using off-the-shelf SMT solvers or by multiple runs of
a linear programming solver. For any counter-example found in the
abstraction, we check the existence of a corresponding violation in
the original system. We demonstrate the efficacy of our technique on a
few dynamical systems examples.

% In this work we address the problem of finding unsafe behaviors with
% given a safety property for a hybrid dynamical system. We treat the
% system as a black box and approach the problem in two distinct steps.
% First, we model the behavior of the system using a piecewise affine
% discrete time model. Such a model is incomplete in nature and is
% computed with respect to the given property. Next, we encode the
% search for falsification as a bounded model checking query and use an
% SMT solver to find a counter example. If found, we check the existence
% of the violation in the original system to ensure reproducibility.

%Time bounded LTL property in
%Models from verification
%model without data
%black box model/data -> structured model