### 2017-04-07 prelims.tex

In this section, we present the model of the hybrid dynamical system
under test, and the background for pwa relational abstractions.

\mypara{The System under test $\System$} is assumed to be a hybrid
dynamical system. Its state is denoted by $\HybridStates \subseteq
\Modes \times \Reals^n$, where $\Modes$ is a finite set of discrete
modes. Let $\Inputs: [0,T] \rightarrow \reals^k$ be the set of input
signals to $\System$ over a finite time horizon $T$.

For the purpose of falsification, $\System$ is viewed as a black-box
equipped with a forward simulation map $\simulate^{\System}:
(\HybridStates, \vu, t) \mapsto \HybridStates$.  A simulation step
computes the unique reachable state $\vx \in \HybridStates$ in time
$t$ from a given state input pair $(\vx,\vu)$.  We formalize this
using the notion of time parameterized reachability relations
$\reach{t, \vu}_\System \subseteq \HybridStates \times \HybridStates$
detailed in~\cite{zutshi2014multiple}.
Hence, $\vy=\simulate^{\System}(\vx,u,t)$ \emph{iff} $\vx \reach{t,u}
\vy$.
% Given the hybrid state space $\vx\in\HybridStates$ and input space $u
% \in\Inputs$ and time step $t$, it computes the unique successor state
% $\vy=\simulate^{\System}(\vx,u,t)$ such that $\vx \reach{t,u} \vy$.

\myipara{System Assumptions:} We require that $\System$ be
forward deterministic. This
requires existence and uniqueness of trajectories over a finite time
horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
the vector field in each hybrid mode and ruling away issues such as
finite escape times~\cite{Meiss/2007/Differential}. From a practical
standpoint, these assumptions do not pose significant restrictions for
the type of problems that we wish to address. Finally, we assume full
observability of the system state, validate our results only against
the $\simulate$ function and {\em do not} assume that an analytic closed-form
representation of the system dynamics is available.

As the relations abstract the solution of a continuous time,
foward deterministic hybrid dynamical
system, they are deterministic, causal and satisfy the semi-group property.

From here on, we drop the subscript $\System$.

\begin{definition}
    The system under test $\System$ evolves in the set of infinite
    hybrid state-space $\HybridStates$ under the effect of inputs $\vu
    \in \Inputs$. Its evolution is defined by a simulation function
    $\simulate^\System : \HybridStates \times \Inputs \Reals^+ \mapsto
    \HybridStates$, where $\simulate^\System(\x,\vu,\Delta_t)$ maps a
    current state $\x$ at time $t$ to a new state $\x'$ at time
    $t+\Delta_t$, under a constant input $\vu$.
\end{definition}

\subsection{Learning Dynamics using Simple Linear Regression}

Best effor falsification works by enumerating the reachability
relations $\reach{}$ within a given budget. Depending upon the
specific instance of the problem, the budget can be prohibhitive,
requiring exponential computational and storage resources. Instead, we
proceed by summarizing these relations by their corresponding affine
maps using linear regression. The learnt analytical forms are
represented as $\vx' = A\vx + B\vu + \delta$, where $A$ and $B$ are
matrices and $\delta$ the estimated error.

In statistics, linear regression is the problem of finding an affine
function or \textit{predictor}, which can `best' summarize the
relationship between the given set of observed input $\x$ and output
$\y$ vectors. The notion of `best' is formally captured using a loss
function which can be non-linear. We use the commonly used loss
function \textit{ordinary least squares} (OLS) for this presentation.

\mypara{Ordinary Least Squares (OLS)}

Let the data set be comprised of $N$ input and output pairs $(\x, y)$,
where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is the case in
the current context of \textit{finding the best fit}, the problem is
over-determined: there are more equations than unknowns. Hence, a
single affine function cannot be found which satisfies the equation
$\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we need
to find the `best' choice for $\vec{a}$ and $b$. This choice is formally
defined using a loss function. For the case of simple linear
regression or OLS, the loss function is the sum of squares of the
errors in prediction.  The task is then to determine the vector of
coefficients $\vec{a}$ and an offset constant $b$, such that the least
square error of the affine predictor is minimized for the given data
set.
\begin{equation}
    \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
\label{reg}
\end{equation}
To ease the presentation, we can rewrite the above as a homogeneous
expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
$\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
\[
    \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
\]
The solution of OLS can be analytically computed as
\[ Ab = (X^TX)^{-1}X^T\y\]
where $X$ is the matrix representing the horizontal stacking of all
$\hat{\x}$.
The details can be found in standard texts on learning and statistics
\cite{friedman2001elements}.

Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
where $\x\in\reals^n$ we can use OLS to approximate its trajectories
at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
$A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
\begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
    global relational model for the system. The data set for the
    regression is a set of pairs $\setof{(\x,\x') | \x' =
    \simulate(\x, \Delta)}$ and can be generated on demand. Another
    set can be generated to compute the error $\delta$ as an interval
    of vectors, where each element is an interval $\delta_i \in
    [\delta_{min}, \delta_{max}]$.

    Affine maps are poor
    approximations for arbitrary non-linear functions. Hence, we use a
    collection of affine maps to approximate the local behaviors (in
    state-space) of the system $\simulate$. This results in a piece
    wise approximation, as described in the next section.

% \mypara{Regression Analysis}

% In statistics, regression is the problem of finding a
% \textit{predictor}, which can suitably predict the relationship
% between the given set of observed input $\x$ and output $\y$ vectors.
% In other words, assuming that $\y$ depends on $\x$, regression
% strategies find either a parameterized or a non-parameterized
% prediction function to explain the dependence. We now discuss simple
% linear regression, which is parametric in nature and searches for an
% affine predictor. It is also called \textit{ordinary least squares}.

% \mypara{Ordinary Least Squares (OLS)}

% Let the data set be comprised of $N$ input and output pairs $(\x,
% y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
% the case in the current context of \textit{finding the best fit}, the
% problem is over-determined; there are more equations than
% unknowns. Hence, a single affine function cannot be found which
% satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
% need to find the `best' choice for $\vec{a}$ and $b$. This is formally
% defined using a loss function. For the case of simple linear
% regression or OLS, the loss function is the sum of squares of the
% errors in prediction.  The task is then to determine the vector of
% coefficients $\vec{a}$ and an offset constant $b$, such that the least
% square error of the affine predictor is minimized for the given data
% set.
% \begin{equation}
%     \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
% \label{reg}
% \end{equation}
% To ease the presentation, we can rewrite the above as a homogeneous
% expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
% $\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
% \[
%     \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
% \]
% The solution of OLS can be analytically computed as
% \[ Ab = (X^TX)^{-1}X^T\y\]
% where $X$ is the matrix representing the horizontal stacking of all
% $\hat{\x}$.
% The details can be found in several texts on learning and statistics
% \cite{friedman2001elements}.

% Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
% where $\x\in\reals^n$ we can use OLS to approximate its trajectories
% at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
% \begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
%     global relational model for the system. The data set for the
%     regression is a set of pairs $\setof{(\x,\x') | \x' =
%     \simulate(\x, \Delta)}$ and can be generated on demand. Another
%     set can be generated to compute the error $\delta$ as an interval
%     of vectors, where each element is an interval $\delta_i \in
%     [\delta_{min}, \delta_{max}]$.
%
%     Affine maps are poor
%     approximations for arbitrary non-linear functions. Hence, we use a
%     collection of affine maps to approximate the local behaviors (in
%     state-space) of the system $\simulate$. This results in a piece
%     wise approximation, as described in the next section.

% \subsection{Relational Abstractions}
% For a hybrid automaton model, they can summarize the continuous
% dynamics of each mode using a binary relation over continuous states.
% The resulting relations are timeless (independent) and hence valid for
% all time as long as the mode invariant is satisfied. The relations
% take the general form of $R(\x,\x') \bowtie 0$, where $\bowtie$
% represents one of the relational operators $=, \ge, \le, <, >$. For
% example, an abstraction that captures the monotonicity with respect to
% time for the differential equation $\dot{x} = 2$ is $x' > x$. The
% abstraction capturing the relation between the set of ODEs: $\dot{x} =
% 2$, $\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.

%%  Similar to timeless relational abstractions, time-aware relational
%%  abstractions \cite{mover2013time} construct binary relations between
%%  the current state $\x$ of the system and any future reachable state
%%  $\x'$. The difference lies with the latter also constructing relations
%%  between the current time $t$ and any future time $t'$. This is
%%  achieved by a case by case analysis of the eigen structure of the
%%  matrix $A$ (of an affine ODE). Separate abstractions are used for the
%%  case of linear systems with constant rate, real eigen values and
%%  complex eigenvalues.

\subsection{Piecewise Affine (PWA) Transition System}

We define the PWA transition system as a transition system
%(ref.~\secref{DiscSys})
$\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$
where each transition $\tau \in \scrT$, is associated with a
transition relation of the form
\[
    \rho_{\tau}(\pVar,\pVar') \subseteq \setof{(\pVar,\pVar')) \;|\;
    g_\tau(\pVar) \land g'_\tau(\pVar') \land f_{\tau}(\pVar, \pVar')}
\]
where $f_{\tau}$ is an affine relation on the pre and post states
$\pVar,\pVar'$, $g_{\tau}$ and $g'_{\tau}$ are affine guards (pre and
post conditions) on the states and $\Theta$ is an affine assertion on
the initial states.

We can now use a PWA transition system $\rho$ to approximate the
behavior of a hybrid dynamical system $\System$ defined by a simulator
$\simulate^{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
form, we can only approximate general hybrid systems.}. Each
transition of $\rho$ represents a discrete time step $\Delta$. Let the
state-space of $\System$ be given by $\x\in\HybridStates^n$, then the
states of $\rho$ are also given by $\x\in\HybridStates^n$. The affine guard
predicates $g_\tau$ are given by a conjunction of hyper-planes and
hence can be represented in matrix form as a polyhedron (defined by
$m$ constraints) in the state-space
\[
g:C\x - \vd \le 0
\]
where $C$ is an $m \times n$ matrix and $\vd$ is a vector of length
$m$. The affine relation $f_\tau$ can be represented using matrix $A$
and an offset vector $\vb$ as $f_\tau \subseteq \setof{(\x,\x') | \x'
= A\x+\vb}$. Finaly, we incorporte an approximate non-deterministic
error term $\delta$, defined as a vector of intervals
$[\delta_{\min},\delta_{\max}]$.
\[
    f_\tau \subseteq \setof{(\x,\x') | \x' = A\x+\vb}.
\]
%defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
%\in f_\tau(\x)}$.

% Given a state vector $\x\in\reals^n$, a guarded affine map
% $\gm:(\guard, \amap)$ defines the discretized consecution rule as a
% pair of an affine guard predicate $\guard:C\x - \vd \le 0$ and an affine map
% $\amap:A\x+\vb$, where $A \in \reals^{n \times n}$, $C \in
% \reals^{m \times n}$ are matrices and $\vb \in \reals^n$, $\vd
% \in \reals^m$ are vectors. A guarded affine map is satisfied if its
% guard is satisfied.

We now formalize the PWA affine transition system for a dynamical system.
\begin{definition}[PWA Transition System]

Given a hybrid dynamical system over a state-space
$\HybridStates^n$, a PWA transition system $\rho$ is given by the
tuple $\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in
\scrT$ are affine transition relations and $\Theta$ is an affine
predicate over $\x$ and the states $\x\in\HybridStates^n$. The
transition relation is then defined by $\scrT$ with $n$ transition
relations as follows

\begin{equation}
    \scrT = \left\{
        \begin{array}{ll}
            g_1(\x) \land g'_1(\x') \implies f_1(\x,\x') \\
            \ldots\\
            g_n(\x) \land g'_n(\x') \implies f_n(\x,\x') \\
        \end{array}
    \right.
\end{equation}

\end{definition}

%where $f_i:\x \mapsto A_i\x + \vb_i + \delta_i$,
%$g_i(\x):C_i\x-\vd_i\le0$, $g'_i(\x'):C'_i\x'-\vd'_i\le0$ and $\x'$
where $g_i$ and $g'_i$ are affine predicates and $f_i$ are affine
relations and $\x'$ denotes the next state of the system.

A PWA model is \textit{deterministic}, iff for every state
$\x\in\HybridStates^n$, a unique guarded affine map is satisfied and
all errors $\delta$ are singletons. However, in practice such a case
rarely exists. A PWA model will be usually non-deterministic, both due
to multiple choice of transition relations and the reachable states at
the $k^{th}$ time step (or after $k\Delta$ units of time) being a set.
Abusing the notation, we denote the set of states reachable in a
single step in $\rho$ by $\x' = \rho(\x)$.

A PWA transition system is \textit{complete}, iff for every state
$\x\in\HybridStates^n$, there exists at least one satisfied
transition relation $\scrT$. If this is not the case, the system can
deadlock, with no further executions. This usually results from
modeling errors, and from here on, we do not consider such
cases.

% Given a dynamical system over a continuous state-space
% $\ContStates\in\reals^n$, a PWA model is a map $\ContStates \mapsto
% \setof{\gm_1, \gm_2, \ldots, \gm_n}$ from the state-space of the
% dynamical system to a finite set of guarded affine maps
% $\setof{\gm_1, \gm_2, \ldots, \gm_n}$. It defines the consecution
% rule by the affine map of the satisfied $\gm$.

% \begin{equation}
%     \pwa = \left\{
%         \begin{array}{ll}
%             \gm_1: \amap_1(\x) &\text{if}\; \guard_1(\x) \\
%             \gm_2: \amap_2(\x) &\text{if}\; \guard_2(\x) \\
%             \ldots & \ldots\\
%             \gm_n: \amap_n(\x) &\text{if}\; \guard_n(\x) \\
%         \end{array}
%     \right.
% \end{equation}

% \end{definition}

% where $\amap_i = A_i\x + \vb_i$, and $\guard_i(\x) \equiv C_i\x-\vd_i\le0$.
% Abusing the notation, we denote the next state computed by the PWA
% model as $\x' = \pwa(\x)$. A PWA model is \textit{deterministic}, iff
% for every state $\x\in\ContStates$, a unique guarded affine map is
% satisfied. A PWA model is \textit{complete}, iff for every state
% $\x\in\ContStates$, there exists at least one satisfied guarded affine
% map $\gm$.