### 2017-04-08 pwarel.tex
.

A discrete abstraction over a continuous space uses relations to
describe the behavior of the system over time. Such relations abstract
the underlying relation between concrete states, and concrete paths
can no longer be constructed directly. Instead, search procedueres
like Counter-Example Guided Refinement must be used to find concrete
paths. In general, search techniques are exponential with respect to
computational resources and memory. As an alternative, we introduce
data driven `enrichment' of abstractions to approximate the concrete
relations underlying the abstract relations. Informally, we use linear
regression to compute affine maps associated with the relations
between two abstract states. We then show how associating the computed
maps with the relations transforms a discrete abstraction into a PWA
transitions system.

\subsection{Graph Abstraction}

The abstraction of the hybrid dynamical system is defined by
paritioning the combiend state-input space $\HybridStates \times
\Inputs$ into hyper-rectangles or cells $C_i \in \C$. The cells are
pairwise disjoint $C_i \cap C_j \neq \emptyset$ iff $i \neq j$ and
their union $\bigcup_{C_i\in\C}C_i$ is the state-input
space $\HybridStates \times \Inputs$. This discrete abstraction over
the continuous space is implicit and defined using a quantization
function $Q_\epsilon$ parameterized by the precision $\epsilon$.
This further defines an equivalence relations such that
\[
    (\vx_i, \vu_i) \equiv (\vx_j, \vu_j) \;\text{iff}\; Q_\epsilon(\vx_i,
    \vu_i) = Q_\epsilon(\vx_j, \vu_j)
\]
In other words, $(\vx_i, \vu_i) \in C_i \; \text{iff} \; C_i =
Q_\epsilon(\vx_i, \vu_i)$.

\begin{definition}[Graph Abstraction]
    The quantization function $Q_\epsilon$ induces a DAG on the discrete
    abstraction given by a graph $G(V, E)$, where each vertex
    corresponds to a cell $V_i = C_i$, and an edge exists between two
    cells $(C_j, C_k) \in E$ iff there exist state-input pairs
    $(\vx,\vu) \in C_j$, $(\vx', \vu) \in C_k$ such that $\vx =
    \simulate(\vx, \vu )$.
\end{definition}

Given a numerical simulator $\simulate$ and an initial abstraction
defined by a quantization function $\quant_\epsilon$ and a time step
$\Delta$, we pick samples and simulate them to explore the abstraction
on the fly to obtain a graph $G$, which has a finite number of
abstract states $C$ (or cells) and edges $(C,C')$ iff $C
\areach{\Delta} C'$.

% \begin{definition}[Abstract State Graph]. Let ∆ > 0 be a fixed time
%     step. The abstract state graph H(∆) for time step ∆ is defined by
%     the set of vertices C, and edges (Ci, Cj ) whenever Ci ∆ Cj . Let
%     C0 denote the collection of initial cells in C, i.e., cells Ci
%     such that Ci ∩ X0 6= ∅. Further, let Cu denote the set of unsafe
%     cells, or cells Cj such that there is a state xj ∈ Cj that reaches
%     an unsafe state y ∈ Xf within time 0 ≤ tj < ∆. The abstraction
% H(∆) is given by D C, ∆ , C0, Cu E .  \end{definition}

\subsection{Enriched Abstraction}

\input{figs/segtraj_rel_enriched.tex}

Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
trajectory segment between the respective cells. The graph abstraction
$G$ only states that there exists a state $\x \in C$ from which the
system can evolve to a future state $\x' \in C'$. To increase the
precision we iteratively refined the abstraction by state
splitting. Instead, we now propose an `enrichment' $G^R$ of $G$ by computing a
set of local relations $R_{(C,C')}(\x,\x')$ for every edge $(C,C')$,
which non-deterministically describe relations between $\x \in C$ and $\x'
\in C'$. This is illustrated in \figref{enriched-edge}.
%(compare with \figref{segtraj}).

The enriched graph $G^R$ captures the underlying local forward
dynamics describing the evolution of the system in each abstract
state. We represent the dynamics using an affine model with an
interval error. Such a model can either be approximated using learning
methods or computed as a sound (over) approximation using reachability
set computation methods. Because we assume black box semantics, we
only present the former. Using regression analysis on the $start$ and
$end$ states of the witnessed trajectory segments between two cells,
we compute an approximate discrete map along with an error estimate.
Moreover, using the simulation function $\simulate$, additional
trajectory segments (or data) can be generated if required. The data
can be separated into a training set and testing set to compute the
map and the error respectively.  The latter case can be explored if
the symbolic dynamics of the system are known. A tool like
\flowstar~\cite{chen2013flow} can be used to find the reachable set
map.

Observe that $G^R$, a directed reachability graph, is rich enough to
search for concrete behaviors in the system. We call it a time
parameterized PWA relational abstraction. It can be interpreted as an
infinite state discrete transition system, and we can use
off-the-shelf bounded model checkers to find concrete violations of a
given safety property, and even other temporal properties.  We now
discuss the background required to present our ideas.

Instead of using a CEGAR like loop (used in~\cite{zutshi2014multiple}),
we use the generated trajectory segments to learn quantitative models
describing the local behavior of the system. These models are defined
by a set of relations $R \subseteq \HybridStates \times \HybridStates$
for each edge of the reachability graph.