### 2017-04-08 implementation.tex
.

The presented technique was prototyped as a tool: PWA-Rel.
We compared it against S3CAM~\cite{zutshi2014multiple} on several academic
benchmarks.

PWA-Rel was implemented in Python. Its different parts comprised of
the (a) abstraction graph generation, (b) regeression analysis and (c)
BMC. S3CAM was used to compute (a). To compute (b), OLS regression
routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
Python module for machine learning. Finally, the The bounded model
checking (c) was done using the SAL model checker~\cite{SAL-SRI},
which used Yices2~\cite{dutertre2014yices} as the SMT solver. A naive
alternative linear programming (LP) based model checker specific to the PWA
relational abstractions was implemented utilizing PyGLPK.

We used S3CAM to discover the abstraction graph $G$, which was then
filtered to remove the nodes from which error nodes were not
reachable. In our experiments we used $1$-relational modeling to
create the transition system from $G$. Using the BMC, we checked for
the given safety property. If a concrete trace was obtained, it was
further checked for reproducibility by sampling and simulating its `neighborhood' with the
$\simulate$ function. We define `neighborhood' as the abstract states
corresponding to the returned concrete trace. Specifically, $100$
samples were uniformly randomly chosen and simulated and tested for
error traces. This procedure can also be extended to obtaining
different discrete sequences or discrete trace sequences from the
model checker by requesting distinct counter-examples.

% abstract state was sampled to check the presence of a counter-example
% in it neighborhood. This can be further extended by obtaining
% different discrete sequences or discrete trace sequences from the
% model checker.

We tabulate our preliminary evaluation in \tabref{res-rel}. We used
the more difficult benchmarks described in \cite{zutshi2014multiple},
including the Van der Pol oscillator, Brusselator, Lorenz attractor
and the Navigation benchmark. PWA-Rel and S3CAM were ran $10$ times
with different seeds and the running times were averaged. We tabulate
both the total time taken by PWA-Rel+SAL and PWA-Rel+LP to compute the
counter-example and compare against an implementation of S3CAM.
Different abstraction parameters are used for S3CAM and PWA-Rel, due
to the differences in which they operate. PWA-Rel requires a finer
abstraction to begni with in order to minimize the modeling errors.
On the other hand, a coarser abstraction as the intital abstraction is
works better for S3CAM due to the expensive refinement procedure
involved.

\begin{table}[!htbp]
\centering
\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
time taken by the BMC engine. The total time in seconds (rounded off
to an integer) is noted under \textbf{PWA-Rel} and \textbf{S3CAM}.
$TO$ signifies time $>5hr$, after which the search was killed.}
\label{tab:res-rel}
\begin{tabular}{@{}llll@{}}
\toprule
    Benchmark & PWA Rel + SAL & PWA Rel + LP & S3CAM\\
\midrule
    Van der Pol ($\prop_3$)   &$0.2$ & $0.34$ & $0.4$\\
    Brusselator               &$0.06$ & $0.06$ & $0.03$\\
    Lorenz                    &$0.4$ & $0.33$ & $0.58$\\
%B.Ball                    &  &   & \\
    Nav ($\prop_P$)           &$TO$ & $28.5$  & $10$ \\%9423 + 4480
    Nav ($\prop_Q$)           &$TO$ & $26$   & $9.1$ \\%8105 + 970
    Nav ($\prop_R$)           &$TO$ & $39.2$  & $33.3$\\
    Nav ($\prop_S$)           &$TO$ & $71$  & $35$\\
    Bouncing Ball             &$Safe$ & $Safe$  & $7.5$\\

\bottomrule
\end{tabular}
\end{table}

% 1m8s, 2m56.184s, 1m31.059s[f], 1m32.665s,  1m45.312s, 1m22.397s, 2m1.451s[f], 1m12.118s, 1m26.596s
% 0.12, 0.06,         0.33       0.13,       0.22,     FAIL,       ,X           0.08        0.15

As apparent form the results, the performance is comparable in all
benchmarks, but the last one: bouncing ball. PWA-Rel determines it to
be `safe' as it can not find a counter-example in the PWA relational
model. Also, apparent froom the results, PWA-Rel is more scalale with
the LP solver and PWA-Rel+SAL times out on most of the benchmarks.
The efficiency of the naive LP solver can be attributed to the careful
selection of the pareters for S3CAM (finer abstraction), resulting in
a small subgraph which can be easily explored for all paths.

However, we need to explore more benchmarks to be conclusive on the
tool's performance on real world examples. Although, PWA-Rel does not
need to refine like S3CAM, and should be more scalable, it is still
constrained by the accuracy of the pwa models. Its performance can be
explained by the difficulty in finding an abstraction
($\quant_\epsilon$), which is fine enough, to result in accurate
prediction models but coarse enough to be manageable by the current
bmc tools.  To obtain a reproducible counter-example from the model
checker (and avoid false positives), the transition system should be
created from models with high accuracy, for which a finer abstraction
is required.  However, the exploration of a finer abstraction is
exponentially more complex and results in a much bigger transition
system, which the current state of the art bounded model checkers
(which use generic SMT solvers) can not handle under reasonable
resources.