### 2017-05-26 abstract.tex
changed abstract to clarify the confusion apparent in reviews

Models of embedded control systems are often so complex that
falsification and testing approaches can scale only by numerically
simulating them, and hence treating the systems as black-boxes. Common
approaches like sampling-based falsification, use numerical
simulations to search for violations of safety properties. However,
this relegates them to the `best effort' category. On the other hand,
\emph{rigorous} verification (and falsification) techniques require
symbolic models and are comparatively less scalable. We try to bridge
this gap by incorporating data-driven models in a sampling based
search. Using an efficient, on-the-fly discrete abstraction
search procedure, we compute models only for selective regions of the state
space, thus `enriching' the abstraction. We then show how the
`enriched' abstraction can be used to guide the search for violations
without refinement.
%We use grid-based discrete abstractions to partition
%the state-space of the black-box hybrid dynamical system. We then use
%an efficient search procedure to find abstract violations of safety
%properties
%Assuming a grid-based implicit abstraction over the
%state-space of the black-box hybrid dynamical system, we enumerate
%abstract violations of the safety property. We then model the system
%behavior along these violations and search for concrete violations.
The automatic falsification proceeds by (a) randomly exploring the
state-space abstraction using numerically computed trajectories and
enumerating abstract violations, (b) modeling the behavior of the
system along these abstract violations by using discrete-time
piecewise affine (PWA) \emph{relational} models, and finally, (c)
encoding the search for a concrete violation as a bounded model
checking (BMC) problem.  The BMC query is then a disjunction of
conjunctions of linear constraints (i.e. combinations of linear
programs). It can be solved using off-the-shelf SMT solvers or by
multiple runs of a linear programming solver. For any counter-example
found in this `enriched' abstraction, we check the existence of a
corresponding violation in the original system. We demonstrate the
efficacy of our technique on a few benchmarks.

% Models of embedded control systems are often so complex that they can
% be treated as effectively black-box systems. Sampling-based methods
% for testing such systems use numerical simulations to search for
% violations of properties (specified in a suitable requirement
% formalism, such as temporal logic). This constrains them to a `best
% effort' category unlike rigorous verification techniques, which
% however require a model amenable to symbolic analysis. This work tries
% to bridge the gap by presenting an automatic falsification technique
% which learns a data-driven abstraction of the black box system before
% searching for a violation. It proceeds by (a) modeling the behavior of
% the system using a piecewise affine (PWA) discrete-time {\em
% relational} model and, (b) encoding the search for a violation in the
% abstract model as a bounded model checking problem. The bounded model
% checking query in this instance can be viewed as a disjunction of
% conjunctions of linear constraints (i.e. linear programs) and can be
% solved directly using off-the-shelf SMT solvers or by multiple runs of
% a linear programming solver. For any counter-example found in the
% abstraction, we check the existence of a corresponding violation in
% the original system. We demonstrate the efficacy of our technique on a
% few dynamical systems examples.

% =========================================================================
% Sampling based falsification techniques for black box hybrid dynamical
% systems use numerical simulations to search for violations of
% properties (specified in temporal logic). This constraints them to a
% `best effort' category unlike rigorous verification techniques, which
% however require a white box model. This work tries to bridge the gap
% by presenting an automatic falsification technique which models the
% black box system before searching for a violation. It proceeds by (a)
% modeling the behavior of the system using a piecewise affine (PWA)
% discrete time model and, (b) encoding the search for a violation as a
% standard bounded model checking problem. The resulting problem is a
% disjunction of linear programs and can be solved directly using
% off-the-shelf SMT solvers or by multiple runs of a linear programming
% solver. If a counter-example is found, the existence of a
% corresponding violation in the original system is checked for
% reproducibility.

% In this work we address the problem of finding unsafe behaviors with
% given a safety property for a hybrid dynamical system. We treat the
% system as a black box and approach the problem in two distinct steps.
% First, we model the behavior of the system using a piecewise affine
% discrete time model. Such a model is incomplete in nature and is
% computed with respect to the given property. Next, we encode the
% search for falsification as a bounded model checking query and use an
% SMT solver to find a counter example. If found, we check the existence
% of the violation in the original system to ensure reproducibility.

%Time bounded LTL property in
%Models from verification
%model without data
%black box model/data -> structured model