### 2017-06-12 appendix.tex

============================================================================
EMSOFT 2017 Reviews for Submission 159
============================================================================

Title: Piece-Wise Relational Models for Falsifying Safety Properties of Hybrid Systems

============================================================================
                            REVIEWER 1
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 2
                       Originality (1-4): 4
               Technical Soundness (1-4): 2
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

1. The title states that the paper addresses "hybrid systems" and the system
   model includes discrete states, but thereafter the paper only treats
   continuous systems. Apparently this is reasonable because of an
   assumption that the model is a "switched-mode dynamical system".

   It would be nice to define exactly what you mean by "switched-mode
   dynamical system". Does it mean that only the controller model makes
   discrete transitions and that the environment is a continuous model with
   no discrete events or discontinuities?

   ============
   The systems under test are assumed to be switched-mode to ease the
   presentation, but it is not a limitation of the presented
   technique. Black box systems which have hybrid dynamics due to
   multiple modes, are dealt in the same fashion.
   [TODO: $K$-refinement to differentiate the dynamics of different modes.
   We have added a benchmark to showcase this.
   [TODO: which one?]
   [TODO: clarify the absence of any controllers]
   [TODO: remove or clarify switched-mode systems]
   [TODO: add a hybrid benchmark]
   ============

   I imagine that the idea is to combine the generated discrete abstraction
   with a discrete controller model to verify both together. It would be
   good to state this clearly and to treat the resulting issues throughout
   the paper. I note that the experimental results do not treat this issue.
   Otherwise, why not change the title and focus the text on what you
   actually did?

   ============
   We will explore this in the future, but the current presentation
   only works with the plant, which is itself a hybrid dynamical
   system.
   [TODO: Add this in the conclusion/future work]
   ============

   (Similar remarks hold at a smaller scale. The discussion often describes
    what "can be" done, e.g., in sections 3.2 and 4.1. I would rather read
    about what "has been" done and what the result was!)

   ============
   [TODO]
   ============

2. Why do you need a bounded model checker to check safety invariants when
   you already enumerate and characterize the reachable states? That is, why
   not just test the invariant on the convex hulls as they are generated?

   My guess is that you later want to restrict the state-space by composing
   your abstraction with a controller model. Is that right? This probably
   seems obvious to you, but please put yourself in the place of someone who
   is new to the work and not an expert in the domain! In any case, this
   idea is not clearly stated, not addressed by the text, and not considered
   in the experimental evaluation.

   ============
   [TODO: I don't understand this comment, but it is related to
   point (iii) of the summary]
   Although we enumerate the abstract space, we require the model
   checker to find a concrete counter-example.
   ============

   Is this why the PWA transition system characterizes states by a valuation
   function v and not simply as tuples? If yes, how will this work? I do not
   understand why you have a valuation function when the system only has one
   variable (x).
   ============
   [TODO]
   ============
============================================================================
                            REVIEWER 2
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 4
          Presentation/Readability (1-4): 3
                       Originality (1-4): 4
               Technical Soundness (1-4): 3
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

- In section 2.1. the authors state that when N>n “a single affine function
cannot be found”. Should the problem be stated in finding the best choice of
a and b or should the problem considering determining the optimal number of
piece wise affine approximations? Can the magnitude N-n give a hint of what is
a good number of PWA / affine approximations? I feel that this could lead to a
nice results which the authors can easily cast in optimization (2).

>> In general, finding the `optimal' number of PWA models can be
>> framed as a multiobjective optimization problem, where a small
>> number of clusters and an accurate affine map is desired. This is a
>> hard problem for arbitrary non-linear systems.
>> Hence, in this paper, we presented a scalable falsification approach
>> with a sub-optimal PWA identification approach.

Can the authors comment or speculate on how the complexity of the
dynamics might influence the required number of affine maps? How does
the non-smoothness and non-differentiability influence the computation
of R(C,C’)? Is there any relationship between the number of
nondifferentiable points and the number of required affine maps?

- Another critical parameter in the framework seems to be the time steps Delta
used in section 4. Can the authors perform a sensitivity analysis to gauge its
impact on overall quality of the results?

>> It is a good idea, and it can have a big affect on the quality of
the results. However, selecting a `good' Delta is not straightforward.

- Since the main goal of this technique is to find violations of safety
properties in the embedded control system, it would prove useful for the reader
if the authors could comment which of the evaluations are considering this
specific goal. Again, this is not a criticism, but I feel that the
investigation in section 6 could be described in a few more sentences to
connect with embedded control.

============================================================================
                            REVIEWER 3
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 3
                       Originality (1-4): 2
               Technical Soundness (1-4): 3
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

The paper presents a method for falsifying safety properties of hybrid systems.
The method first constructs a relational abstraction of the system (which is
considered as a black box) from a set of simulation traces. The relations in
the abstraction are affine with uncertainty. The approach is quite interesting,
however the examples used for demonstration are unfortunately not black boxes,
although these examples are treated as black boxes, since it would be of great
interest to show its usefulness for complex models for which traditional
verification tools cannot be applied. A comparison with S3CAM is interesting,
but is less informative than a comparison with simulation-based validation
tools.
I also recommend using relations more complex than affine, such as polynomial,
for which Z3 can be used (nonlinear relations may reduce the number of
hyper-rectangles).

The method is of heuristic nature and thus convincing experimental results
would be needed. The paper compares with previous work (S3CAM) using a similar
approach. This approach has been tested on more real-life examples from
automative control. I am wondering why there are no comparisons on these
examples.

============================================================================
                            REVIEWER 4
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 2
                       Originality (1-4): 3
               Technical Soundness (1-4): 2
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

Although the author claims that the proposed technique can effectively find a
counter-example for a black-box model, however, the experiment result does not
reflect that statement. Most of the benchmarks are nonlinear continuous systems
that are not black-box. I strongly suggest the author evaluate the proposed
approach on some black-box systems such as industrial Simulink models.

Using OLS does not always result in finding the best fit. For example, if some
data points are excessively small or large compared to the rest of data, they
will have disproportionately large effects on the resulting constant. In this
case, using other techniques like ridge regression or lasso regression may
yield a better result.

>> As the choice of loss function and type of regression analysis
>> depends on the problem at hand, we show our methodology using the
>> simplest one. In practice, sophesticated statistical tests can be used
>> to find the a good analysis.

============================================================================
                            REVIEWER 5
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 3
                       Originality (1-4): 3
               Technical Soundness (1-4): 3
Account of prior work and references (1-3): 3

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

I suggest leading with the Van Der Pol oscillator to help ignorant
readers such as myself get a handle on what it is you're actually able
to do, then launch into the formalism.

>> We have made the Van Der Pol oscillator as a running example to improve
>> the presentation.