### 2017-06-12 implementation.tex

The presented technique was prototyped as a tool: PW-Rel.  We
compared it against S3CAM~\cite{zutshi2014multiple} on several
benchmarks. Below, we present the details of the implementation and
the benchmarks used.

\subsection{Implementation}

PW-Rel is implemented in Python. It uses S3CAM to generate the graph
abstraction and regression routines from
Scikit-learn~\cite{pedregosa2011scikit} to compute
$\RTSAbstraction{\epsilon}{\System}$ of the given system $\System$.
It then constructs relational abstraction and uses a bounded model
checker to find a counter-example to the given property in
$\RTSAbstraction{\epsilon}{\System}$. Finally, using simulations, it
confirms if the counter-example is valid for $\System$, in which case
a successful falsification run is concluded.

To check the transition system, PW-Rel can either invoke symbolic
bounded model checkers or explicit model checkers. For symbolic BMC,
the off-the-shelf SAL model checker~\cite{SAL-SRI} is used. It is
configured to encode the problem as a single query for a given depth
$k$, where $k$ is the bound on the length of the counter-example. It
was computed by the length of the abstract counter-examples found. SAL
uses the SMT solver Yices2~\cite{dutertre2014yices} to solve the
encoded SMT constraints.

% Two symbolic BMC were tried to compare scalability; the off-the-shelf
% SAL model checker~\cite{SAL-SRI} and SatExBMC. Both model checkers
% were configured to encode the problem as a single query for the given
% depth. SAL uses the SMT solver Yices2~\cite{dutertre2014yices}, while
% SatExBMC uses SatEx~\cite{shoukry2017smc}. SatEx differs from Yices2
% (and other sound SMT solvers) in two aspects, (a) it is a lazy SMT
% solver and (b) uses approximate convex optimization routines (via
% CPLEX). Although (b) makes it unsound, it is not a concern for our
% falsification procedure as it is based off approximate models.

Explicit model checking was implemented out using a naive DFS
(depth-first-exploration) strategy, wherein, each abstract path is
enumerated and checked for feasibility using a an optimization
routine. Interfaces to various solvers were implemented, including GLPK
and Ipopt. If feasible, the returned concrete counter-example is
simulated and checked for existence in $\System$. Also, PW-Rel
extracts the bounds on the initial parameters by formulating multiple
optimization problems. The bounds can be regarded as an over-appr

% If the existence check fails, multiple
% parameter values can be sampled within the determined bounds to search
% for counter-examples in the neighborhood of the returned concrete
% counter-example.

% PW-Rel uses S3CAM to discover the abstraction graph $G$, which is
% then filtered to remove the nodes from which error nodes are not
% reachable. In all our experiments we used $1$-relational modeling to
% create the transition system from $G$. This was determined to be a
% good trade-off between longer simulations and increase in precision,
% with respect to the mentioned benchmarks.
% \todo{$1$-relational modeling was used}

Using BMC, we checked for the given safety property. If a concrete
trace was obtained, it was further checked for reproducibility by
sampling and simulating its `neighborhood' with the $\simulate$
function. We define `neighborhood' as the abstract states
corresponding to the returned concrete trace.  Specifically, $100$
samples were uniformly randomly chosen and simulated and tested for
error traces. This procedure can also be extended to obtaining
different discrete sequences or discrete trace sequences from the
model checker by requesting distinct counter-examples.

% abstract state was sampled to check the presence of a counter-example
% in it neighborhood. This can be further extended by obtaining
% different discrete sequences or discrete trace sequences from the
% model checker.

\mypara{The search parameters} for PW-Rel include the parameters of
S3CAM: $N$, $\epsilon$ and $\Delta$. Additionally, they also include
the maximum error budget $\delta_{max}$ for OLS and the maximum length
of segmented trajectory for building $k$-relational models.  The
effects of $N$, $\epsilon$ and $\Delta$ on S3CAM's performance were
detailed in~\cite{zutshi2014multiple}. However, they also have an
effect on relational modeling. A finer grid with small cells produces
more accurate models than a coarser grid with large cells. Similarly,
small time length trajectory segments result in more accurate
modeling.

%An increase in the number of samples can increases the
%accuracy of regression.

% \paragraph{Maximum Model Error ($\boldsymbol{\delta}_{\bold{max}}$) .}
% Given a $\delta_{max}$, we keep increasing $k$ during the
% $k$-relational modeling process till $\delta_i \le \delta_{max}$ is
% satisfied. A $k_{max}$ is introduced to bound the longest
% segmented trajectory which can be considered.

% \paragraph{Bounded k relations ($\boldsymbol{k}_{\bold{max}}$.) }

\subsection{Evaluation}

We benchmarked our implementation and compared it with S3CAM.  The
preliminary results are tabulated in~\tabref{res-rel}. We used few of
the benchmarks described in \cite{zutshi2014multiple}, including the
Van der Pol oscillator, Brusselator, Lorenz attractor and the
Navigation benchmark.
%We also tried PW-Rel on the insulin
%benchmark~\cite{chenformal} and lane change maneuver
%benchmark~\cite{althoff2012reachability}.
PW-Rel was compared with
S3CAM by running both $10$ times with different seeds (as both use
randomized algorithms), and their running times were averaged. PW-Rel
was evaluated with both the symbolic BMC SAL (PW-Rel+SAL) and the
explicit BMC (PW-Rel+OPT). %\todo{SatEx was also tried, but ...}.
An affine kernel was used for modeling and the linear programming
solver GLPK was used for checking feasibility of the constraints.
We also tried polynomial kernels, and used Z3 SMT
solver~\cite{DeMoura+Bjorner/08/Z3} to check the symbolic BMC.
Unfortunately, Z3 does not scale well on polynomial constraints and
consistently timed out. For explicit BMC, several non-linear
optimization engines were tried, but the results were inconsistent due
to sensitivity towards the initial guess.

We tabulate both the total time taken by PW-Rel+SAL and PW-Rel+OPT to
compute the counter-example and compare against an implementation of
S3CAM.  Different abstraction parameters are used for S3CAM and
PW-Rel, due to the differences in which they operate. PW-Rel
requires a finer abstraction to begin with in order to minimize the
modeling errors. On the other hand, a coarser abstraction as the
initial abstraction works better for S3CAM due to the expensive
refinement procedure involved.~\footnote{In all our experiments we
used $1$-relational modeling to create the transition system.
This was determined to be a good trade-off between longer simulations
and increase in precision, with respect to the mentioned benchmarks.}

\begin{table}[!htbp]
\centering
\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
time taken by the BMC engine. The total time in minutes is noted under
    \textbf{PW-Rel+BMC solver} and \textbf{S3CAM}.
$TO$ signifies time $>5hr$, after which the search was killed.}
\label{tab:res-rel}
\begin{tabular}{@{}llll@{}}
\toprule
    Benchmark & PW-Rel + SAL & PW-Rel + OPT & S3CAM\\
\midrule
    Van der Pol ($\prop_3$)   &$0.2$ & $0.34$ & $0.4$\\
    Brusselator               &$0.06$ & $0.06$ & $0.03$\\
    Lorenz                    &$0.4$ & $0.33$ & $0.58$\\
%B.Ball                    &  &   & \\
    Nav ($\prop_P$)           &$TO$ & $28.5$  & $10$ \\%9423 + 4480
    Nav ($\prop_Q$)           &$TO$ & $26$   & $9.1$ \\%8105 + 970
    Nav ($\prop_R$)           &$TO$ & $39.2$  & $33.3$\\
    Nav ($\prop_S$)           &$TO$ & $71$  & $35$\\
    Bouncing Ball             &$Safe$ & $Safe$  & $7.5$\\

    %insulin-st-1             &$Safe$ & $Safe$  & $0.0$\\
    %insulin-st-2             &$Safe$ & $Safe$  & $0.0$\\
    %lane change              &$Safe$ & $Safe$  & $0.0$\\
    %HVAC                     &$Safe$ & $Safe$  & $0.0$\\

\bottomrule
\end{tabular}
\end{table}

% 1m8s, 2m56.184s, 1m31.059s[f], 1m32.665s,  1m45.312s, 1m22.397s, 2m1.451s[f], 1m12.118s, 1m26.596s
% 0.12, 0.06,         0.33       0.13,       0.22,     FAIL,       ,X           0.08        0.15

In all the runs of both the tools, a falsification was found. As
apparent form the results, the performance is comparable in all
benchmarks, but the last benchmark: bouncing ball. PW-Rel determines
it to be `safe' as it can not find a counter-example in the piecewise
relational model. Also, apparent from the results, PW-Rel is more
scalable with the LP solver and PW-Rel+SAL times out on most of the
benchmarks.  The efficiency of the naive LP solver can be attributed
to the careful selection of the parameters for S3CAM (finer
abstraction), resulting in a small subgraph which can be easily
explored for all paths.

However, we need to explore more benchmarks to be conclusive on the
tool's performance on real world examples. Although, PW-Rel does not
need to refine like S3CAM, and should be more scalable, it is still
constrained by the accuracy of the piecewise models. Its performance can be
explained by the difficulty in finding an abstraction
($\quant_\epsilon$), which is fine enough, to result in accurate
prediction models but coarse enough to be manageable by the current
BMC tools.  To obtain a reproducible counter-example from the model
checker (and avoid false positives), the transition system should be
created from models with high accuracy, for which a finer abstraction
is required.  However, the exploration of a finer abstraction is
exponentially more complex and results in a much bigger transition
system, which the current state of the art bounded model checkers
(which use generic SMT solvers) can not handle under reasonable
resources.

\subsection{Reasons for Failure}

The approach can fail to find a counter-example, even when it exists,
in one of three ways. As it is dependant on S3CAM to find abstract
paths, it can fail when (a) S3CAM fails to find abstract
counter-examples. We can remedy it by increasing search budgets and/or
restarting. Even when abstract counter-examples are found, they may
not be \emph{concretizable}. In such a case, the (b) BMC search fails.
The failure can be attributed to either spurious abstract
counter-examples or a poorly estimated model. The former can be
addressed by restarting but the latter requires that the maximum model
error $\delta_{max}$ be decreased. Finally, the (c) counter-example
from the BMC might not be reproducible in the black box system. This
happens if the relational model is not precise enough.